Novelty Detection Using Graphical Models for Semantic Room Classification

### 2011-05-17 semantic-nd/article.tex
Introduction.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
There has been several efforts in the area of Artificial Intelligence and Robotics
in creating robots that are able to interact with humans and their environments.
One of the existing problems is a reliable high-level localization method that can
be deployed into new and unknown environments or where its impossible to access a
previously labelled map of the area. Nonetheless the need of reliability and ability
to deal with new cases also arises due to dynamic changes on the environment through
time and with human action.
This article focus on the mapping, using \emph{semantic data}, of indoor
environments such as houses and offices to room categories such as kitchen,
corridor, office.

Due to the unknown and dynamic nature of the operating environment the
mapping methods should be resilient to new room categories the agent has never seen
before. This is: perform novelty detection on the concepts the agent is aware of.

Dora (CogX: Dora) was used as a base system where to implement the
presented novelty detection system as it already possesses a developed architecture
oriented on dealing with probabilistic and semantic information and performing
reasoning over it.
Implementing novelty detection capabilities on an agent such as Dora, would allow to
to increase reliability and allow the creation of self-extending behaviours.

Its with a final interest on integrating the developed methods on Dora that this
initial work focus on the novelty detection using probabilistic graphical models
that model the conceptual knowledge and information handled by Dora.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

There has been several efforts in the areas of artificial intelligence and robotics
in creating robots that are able to interact with humans and their environments.
One of the important aspects is to endow those robots with a deeper understanding
of human environments, not just for the purpose of navigation and obstacle avoidance, 
but also in terms of human semantics and functionality. An important problem
in creating reliable representations of space for robots that are to be deployed 
in new and unknown realistic environments is to be able to automatically 
identify gaps in robot's knowledge and act in order to fill those gaps.

This article addresses the problem of \emph{novelty detection} within the context
of semantic mapping i.e. generating maps containing \emph{semantic information} 
about indoor environments, such as homes or offices. In that context, the ability 
to detect that the observations result from a semantic concept unknown to the robot, 
and cannot be explained by one of its models, is crucial for generating fully autonomous
and reliable behavior. In order to be robust, the robot must identify 
novel concepts and instead of making a costly error, refrain from the decision
and initiate learning. In particular, we address the problem of novelty detection for
room categorization i.e. detecting whether the area in the environment identified as a
separate room can be assigned one of the semantic labels that the robot knows (e.g. as \emph{a kitchen} 
or \emph{an office}) or belongs to a yet unknown semantic category.

The novelty detection algorithm is implemented within a cognitive robot Dora the Explorer,
which already possesses a developed architecture based on proababilistic graphical models 
oriented towards dealing with uncertain semantic information and reasoning about it~\cite{ijcai}.
One of the major problems in using the probabilistic models of the environments for novelty
detection is selecting the optimal threshold above which the test sample is considered
novel. This problem becomes more difficult when, as in case of dora, the representation
grows as the robot explores the environment. Methods are required to find
optimal parameters given the current structure of the model. To this end, this work 
studies methods for novelty threshold selection using probabilistic graphical models.

The rest of this paper is structured as follows. First, we briefly review the related work on
room categorization and novelty detection (Section~\ref{}). Then, we give an outline of the 
architecture of Dora and the structure of the conceptual map representing the semantic information
(Section~\ref{}). Next, we discuss the methods for novelty detection (Section~\ref{}) 
and present results of our prelimianary experiments (Section~\ref{}). This paper concludes with a 
summary in Section~\ref{}.
